Supplement: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
capitalize titles

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string methods provided by Python --
and recover many of the higher-level building blocks that have made these
other system successful.
Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \systemEnding.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
At first, \system cannot programs for most of the tasks.
After three wake/sleep iterations,
it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

>>>>

\subsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string methods provided by Python --
and recover many of the higher-level building blocks that have made these
other system successful.
Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \systemEnding.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
At first, \system cannot programs for most of the tasks.
After three wake/sleep iterations,
it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

### 2018-02-09 supplement.tex
capitalize titles

\section{An illustration of the wake-sleep cycle}
Below we diagram the wake/sleep cycles employed by our algorithm. At each stage of the cycle,
we have shaded the observed variables in gray and left the unobserved variables white.
Black lines correspond to a connection from the top-down
generative model, while red lines correspond to connections from the bottom-up recognition model.

@@

\section{Neural recognition model architecture}

@@

\subsection{Structure learning}

@@

\section{Initial domain-specific languages}
\subsection{Boolean circuits}
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
    nand & $\text{bool}\to \text{bool}\to \text{bool}$ \\  \hline
\end{tabular}
\end{center}

\subsection{Symbolic regression}
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
  + & $\mathbb{R}\to \mathbb{R}\to \mathbb{R}$ \\
  *&  $\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
  real&$\mathbb{R}$\\\hline
\end{tabular}
\end{center}

\subsection{String editing}
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
  0&$\mathbb{Z}$\\
  +1&$\mathbb{Z}\to \mathbb{Z}$\\
  -1&$\mathbb{Z}\to \mathbb{Z}$\\
  split&$\text{char}\to \text{string}\to \left[\text{string} \right]$\\
  join&$\text{string}\to \left[\text{string} \right]\to \text{string}$\\
  map&$(\text{string}\to \text{string})$\\
  &$\qquad\quad\to\left[\text{string} \right]\to \left[\text{string} \right]$\\
  nth&$\mathbb{Z}\to \left[\text{string} \right]\to \text{string}$\\
  slice&$\mathbb{Z}\to \mathbb{Z}\to \text{string}\to \text{string}$\\
  length&$\text{string}\to \mathbb{Z}$\\
  chr->str&$\text{char}\to \text{string}$\\
  trim&$\text{string}\to \text{string}$\\
  upper&$\text{string}\to \text{string}$\\
  lower&$\text{string}\to \text{string}$\\
  capitalize&$\text{string}\to \text{string}$\\\hline
\end{tabular}
\end{center}
We also provide the constant empty string which has type string and also provide
constant characters which have the type char.

@@

\subsection{List functions}
The lower section with seven primitives designates those given to the
\emph{rich} DSL variant and not the \emph{base} variant. The rich variant is
not supplied the \texttt{if} primitive, because its primary intention is to
enable the learning of the lower five concepts.
Greek letters represent polymorphic types.
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
    empty & [$\alpha$] \\
    singleton & $\alpha \to$ [$\alpha$] \\
    concat & [$\alpha$] $\to$ [$\alpha$] $\to$ [$\alpha$] \\
    mapi & $(\mathbb{Z} \to \alpha \to \beta)$ \\
    & \qquad\qquad$\to$ [$\alpha$] $\to$ [$\beta$] \\
    reducei & $(\beta \to \mathbb{Z} \to \alpha \to \beta)$\\
    & \qquad\qquad$\to \beta \to$ [$\alpha$] $\to \beta$ \\
    \hline
    if & bool $\to \alpha \to \alpha \to \alpha$ \\
    \hline
    true & bool \\
    not & bool $\to$ bool \\
%UNUSED+UNNECESSARY:    and & bool $\to$ bool $\to$ bool \\
%UNUSED+UNNECESSARY:    or & bool $\to$ bool $\to$ bool \\
    0 $\cdots$ 5 & $\mathbb{Z}$ \\
    + & $\mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ \\
    * & $\mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ \\
    negate & $\mathbb{Z} \to \mathbb{Z}$ \\
    mod & $\mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ \\
    eq? & $\mathbb{Z} \to \mathbb{Z} \to$ bool \\
    gt? & $\mathbb{Z} \to \mathbb{Z} \to$ bool \\
    is-prime & $\mathbb{Z} \to$ bool \\
    is-square & $\mathbb{Z} \to$ bool \\
    range & $\mathbb{Z} \to$ [$\mathbb{Z}$] \\
    sort & [$\mathbb{Z}$] $\to$ [$\mathbb{Z}$] \\
    \hline
    sum & [$\mathbb{Z}$] $\to \mathbb{Z}$ \\
    reverse & [$\alpha$] $\to$ [$\alpha$] \\
    all & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ bool \\
    any & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ bool \\
    index & $\mathbb{Z} \to$ [$\alpha$] $\to \alpha$ \\
    filter & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ [$\alpha$] \\
    slice & $\mathbb{Z} \to \mathbb{Z} \to$ [$\alpha$] $\to$ [$\alpha$] \\
  \hline
\end{tabular}
\end{center}

@@

\section{Hyperparameters \& Implementation details}
We set structure penalty $\lambda = 1$ (Eq.~5 of the main paper) and
smoothness parameter $\alpha = 10$ (Eq.~6 of the main paper)
for all experiments except for string editing.
For string editing we found that more strongly regularizing
the DSL was necessary,
so we set $\lambda = 5$.
We believe this is because
we have many more tasks for string editing than we do for the other domains.
The frontier size used for Boolean circuits was $5000$,
for symbolic regression we used $500$,
for string editing we used $10^5$,
and for list functions we used $10^4$.

@@

\section{Why not the ELBO bound?}

@@

\section{Example program induction tasks}

\subsection{String editing}

>>>>

\section{An Illustration of the Wake-Sleep Cycle}
Below we diagram the wake/sleep cycles employed by our algorithm. At each stage of the cycle,
we have shaded the observed variables in gray and left the unobserved variables white.
Black lines correspond to a connection from the top-down
generative model, while red lines correspond to connections from the bottom-up recognition model.

@@

\section{Neural Recognition Model Architecture}

@@

\subsection{Structure Learning}

@@

\section{Initial Domain-Specific Languages}
\subsection{Boolean Circuits}
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
    nand & $\text{bool}\to \text{bool}\to \text{bool}$ \\  \hline
\end{tabular}
\end{center}

\subsection{Symbolic Regression}
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
  + & $\mathbb{R}\to \mathbb{R}\to \mathbb{R}$ \\
  *&  $\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
  real&$\mathbb{R}$\\\hline
\end{tabular}
\end{center}

\subsection{String Editing}
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
  0&$\mathbb{Z}$\\
  +1&$\mathbb{Z}\to \mathbb{Z}$\\
  -1&$\mathbb{Z}\to \mathbb{Z}$\\
  split&$\text{char}\to \text{string}\to \left[\text{string} \right]$\\
  join&$\text{string}\to \left[\text{string} \right]\to \text{string}$\\
  map&$(\text{string}\to \text{string})$\\
  &$\qquad\quad\to\left[\text{string} \right]\to \left[\text{string} \right]$\\
  nth&$\mathbb{Z}\to \left[\text{string} \right]\to \text{string}$\\
  slice&$\mathbb{Z}\to \mathbb{Z}\to \text{string}\to \text{string}$\\
  length&$\text{string}\to \mathbb{Z}$\\
  chr->str&$\text{char}\to \text{string}$\\
  trim&$\text{string}\to \text{string}$\\
  upper&$\text{string}\to \text{string}$\\
  lower&$\text{string}\to \text{string}$\\
  capitalize&$\text{string}\to \text{string}$\\\hline
\end{tabular}
\end{center}
We also provide the constant empty string which has type string and also provide
constant characters which have the type char.

@@

\subsection{List Functions}
The lower section with seven primitives designates those given to the
\emph{rich} DSL variant and not the \emph{base} variant. The rich variant is
not supplied the \texttt{if} primitive, because its primary intention is to
enable the learning of the lower five concepts.
Greek letters represent polymorphic types.
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
    empty & [$\alpha$] \\
    singleton & $\alpha \to$ [$\alpha$] \\
    concat & [$\alpha$] $\to$ [$\alpha$] $\to$ [$\alpha$] \\
    mapi & $(\mathbb{Z} \to \alpha \to \beta)$ \\
    & \qquad\qquad$\to$ [$\alpha$] $\to$ [$\beta$] \\
    reducei & $(\beta \to \mathbb{Z} \to \alpha \to \beta)$\\
    & \qquad\qquad$\to \beta \to$ [$\alpha$] $\to \beta$ \\
    \hline
    if & bool $\to \alpha \to \alpha \to \alpha$ \\
    \hline
    true & bool \\
    not & bool $\to$ bool \\
%UNUSED+UNNECESSARY:    and & bool $\to$ bool $\to$ bool \\
%UNUSED+UNNECESSARY:    or & bool $\to$ bool $\to$ bool \\
    0 $\cdots$ 5 & $\mathbb{Z}$ \\
    + & $\mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ \\
    * & $\mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ \\
    negate & $\mathbb{Z} \to \mathbb{Z}$ \\
    mod & $\mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ \\
    eq? & $\mathbb{Z} \to \mathbb{Z} \to$ bool \\
    gt? & $\mathbb{Z} \to \mathbb{Z} \to$ bool \\
    is-prime & $\mathbb{Z} \to$ bool \\
    is-square & $\mathbb{Z} \to$ bool \\
    range & $\mathbb{Z} \to$ [$\mathbb{Z}$] \\
    sort & [$\mathbb{Z}$] $\to$ [$\mathbb{Z}$] \\
    \hline
    sum & [$\mathbb{Z}$] $\to \mathbb{Z}$ \\
    reverse & [$\alpha$] $\to$ [$\alpha$] \\
    all & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ bool \\
    any & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ bool \\
    index & $\mathbb{Z} \to$ [$\alpha$] $\to \alpha$ \\
    filter & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ [$\alpha$] \\
    slice & $\mathbb{Z} \to \mathbb{Z} \to$ [$\alpha$] $\to$ [$\alpha$] \\
  \hline
\end{tabular}
\end{center}

@@

\section{Hyperparameters \& Implementation Details}
We set structure penalty $\lambda = 1$ (Eq.~5 of the main paper) and
smoothness parameter $\alpha = 10$ (Eq.~6 of the main paper)
for all experiments except for string editing.
For string editing we found that more strongly regularizing
the DSL was necessary,
so we set $\lambda = 5$.
We believe this is because
we have many more tasks for string editing than we do for the other domains.
The frontier size used for Boolean circuits was $5000$,
for symbolic regression we used $500$,
for string editing we used $10^5$,
and for list functions we used $10^4$.

@@

\section{Why not the ELBO Bound?}

@@

\section{Example Program Induction Tasks}

\subsection{String Editing}
